---
title: "|IS 643 CURRENT TOPICS IN DATA ANALYTICS - PROJECT 2\n|
  \ Data Analytics  \n"
author: "Author:MUSA T GANIYU"
date: Feb. 25th, 2017
output:
  html_document:
    fig_caption: yes
    fig_height: 9
    fig_width: 9
    highlight: tango
    keep_md: yes
    theme: cerulean
    toc: yes
    toc_depth: 2
    toc_float:
      collapsed: yes
      number_sections: yes
      smooth_scroll: no
  pdf_document:
    toc: yes
    toc_depth: '2'
---

* Item-Based COllarborative Filtering, according to Kitts,Freed and Vrieze 2000, is a model-based approach which produces recoommendation that is based on the relationship between the items inferred from the rating matrix.

* Kindly load the libraries below (Or install if your dont have it already)

```{r}
options(warn=-1)

suppressMessages(library(recommenderlab))
suppressMessages(library(knitr))
suppressMessages(require(Amelia))
#suppressMessages(library(ggthemes))
suppressMessages(library(ggplot2))
suppressMessages(library(plyr))
suppressMessages(library(plotly))


```

* We will make use of MovieLense data came with recommenderLab library. 

#DATA
Let's also check the structure as well as the dimension of the data.

```{R}

data(MovieLense)
str(MovieLense)
dim(MovieLense)
```

We want to delve more into recommendation which generate recommendation solely on the population.

#THE ANALYSIS
```{R}
movies <- Recommender(MovieLense[1:900], method="POPULAR")
movies
names(getModel(movies))
```

Here are the topN prediction of the recommendation. 

```{R}
getModel(movies)$topN

recom <- predict(movies, MovieLense[901:903], n=10)
recom
as(recom, "list")

```

Best Five (5) for 3 users

```{R}
recom2 <- bestN(recom, n=5)
as(recom2, "list")
```

Rating matrix with inclusion of original user.
```{R}
recom3 <- predict(movies, MovieLense[901:903], type="ratingMatrix")
as(recom3, "matrix")[, 1:10]


```


We have here the top 1, 3, 5, 10, 15 and 20 recommendation lists to evaluate the recommeder using the "popular"" method by obtaining their confusion matrix and avg results.

```{R}


eval2 <- evaluationScheme(MovieLense[1:900], method="cross", k=4, given=3, goodRating=5)
results <- evaluate(eval2, method = "POPULAR", type="topNList", n=c(1,3,5,10,15,20))

confusionmatrix <- getConfusionMatrix(results)[[1]]
kable(confusionmatrix)

kable(avg(results))

```

#PLOTS

A histogram discribtion.
```{R}

hist(getRatings(recom3), breaks = 100)
```

A plot showing precisio-recall.

```{R}
plot(results, "prec/rec", annotate=TRUE)

```







```{R}

data2 <- read.csv("C:/Users/mascot/Documents/GitHub/MSDA-643/movielense.csv", sep=",", header=TRUE)
```

```{r}
library('Matrix')
library('irlba')
library('reshape')
#suppressWarnings(suppressMessages(library('repmis')))

head(data2)
dim(data2)
str(data2)
names(data2)
newdata <- data2[, c(1,2,3)]
newdata2 <- data2[, c(1,2,3,6)]
names(newdata)

```

```{R}
pca <- prcomp(newdata, scale=TRUE)

decomp <- svd(newdata)
names(decomp)
decomp$d <- diag(decomp$d)
pred <- decomp$u %*% decomp$d %*% t(decomp$v)

newdata[1:10, ]
pred[1:10, 1:3]



decomp2 <- irlba(newdata, nu=2, nv=2)
decomp2$d <- diag(decomp2$d)
pred2 <- decomp2$u %*% decomp2$d %*% t(decomp2$v)

newdata[1:10, ]
pred2[1:10, 1:3]
#B <- dcast(newdata, value.var="rating", fill=0,fun.aggregate=mean)

#movie_sparse <- dcast(newdata2)
#S = irlba(data2, nu=3, nv=3)
#L = sparseMatrix(i=movie[[1]], j=movie[[2]], movie=movie[[3]])


```



```{R}


newdata3 <- as(data2, "binaryRatingMatrix")


#abo <- signature(data2="ratingMatrix")
#class(abo)
#users_sim <- similarity(abo[ , ], method="cosine", which="title")

#as.matrix(users_sim)

Vec_rating <- as.vector(newdata)
head(unique(Vec_rating))

Vec_rating <- Vec_rating[Vec_rating !=0]
Vec_rating<- factor(Vec_rating)
Vec_rating<-as.data.frame(Vec_rating)

#qplot(1,Vec_rating, geom="boxplot")
ggplot(Vec_rating) + ggtitle("Distibution of the rating")

```




```{R}
library('Matrix')
library('irlba')
library('reshape')
library(recommenderlab)
movie_rating_avg <- colMeans(newdata)
movie_rating_avg <- as.data.frame(movie_rating_avg)

m <- ggplot(movie_rating_avg, aes(x ="rating"))
m + stat_count(binwidth = 0.1) + scale_y_sqrt() + ggtitle("Distribution of the average movie rating")



ggplot(movie_rating_avg) + aes(x ="rating")+ stat_count(binwidth = 0.1) + ggtitle("Distribution of the average movie rating")


```



```{R}

matx <- as.matrix(newdata)
image(matx[1:10, 1:3], main = "Heatmap of the first rows and columns")

```



